---
title: "My Document Title"
output: html_document
---

# Triangular distribution

Prepare the document: Delete all the variables that were stored before.
```{r}
#clear all variables :
  rm(list = ls())
```


Create table with all relevant information. Our informations were taken from different sources. The data was given in probability distribution. In order to simplify the code and to get a point estimate, we took of the distributions the mean for normal distributions and the mode for triangular distributions. 

The sources were the following:
- Petterson et al., 2016 (Triangular distributions)
- Schönning et al., 2006 (Normal distributions)

In order to find the arithmetic mean, we used the following equation for the trinagnular distribution:
$$
arithmetic\;mean = \frac{min + mode + max}{3}
$$
For the arithmetic mean of the normal distribution, we use the formula.
$$
arithmetic\;mean\;=\;e^{\mu+\frac{\sigma^2}{2}}
$$

This informations can be updated according to your own sources. 
```{r}
df <- data.frame(
  duration_excretion = c((3+7+12)/3, (3+7+12)/3,exp(1.6 + 1.25^2/2), (15+34+42)/3, (10+15+50)/3, exp(3.6 + 0.2^2/2), (5+10+30)/3, exp(2.0 + 0.85^2/2), exp(4.5 + 0.7^2/2), exp(2.1 + 0.25^2/2)),
  excretion_density = c((10^8+10^(10)+10^(12))/3, (10^8+10^(10)+10^(12))/3, exp(20.7 + 2.3^2/2), (10^4+10^(6)+10^(10))/3, (10^6+10^(7.5)+10^9)/3, exp(13.8 + 2.3^2/2), (10^6+10^(7)+10^9)/3, exp(17.3 + 0.6^2/2), exp(15.0 + 1.7^2/2), exp(5.8 + 1.2^2/2)),  
  incidence_rate = c(0.88/100, 1.28/100,1200/100000, 2.4/100, 0.88/100, 500/100000, 1.28/100, 200/100000, 1100/100000, 30/100000),  
  row.names = c("adeno", "rota_triangular", "rota_normal", "campylo", "salmonella_triangular", "salmonella_normal", "crypto_triangular", "crypto_normal", "giardia", "EHEC")  
)
print(df)
```


Define the other parameters:
- X: Fictive population size. This parameter can be adapted according to population. For higher population samples, the variability of the results will be lower.
- rate: This value for the excretion has been taken as a mean value from the studies *To-Figueras et al., 2000*, *Balcells Gorina, 1989* and *Schouw et al, 2002*
- pathogens: In this list, we store the name of all the pathogens we listed in the table above
```{r}
X <- 500000             # fictive population size  
rate <- 250             # [g/cap/day] 

pathogens <- colnames(df)
```


## Production of feces

In population X, $incidence \times X$ persons are infected and $(1 - incidence) \times X$ persons are not infected during the whole year. When a person is not infected, he or she does not excrete the pathogen.


Production of not infected feaces [g/year] is devided into two subparts. For once, the production of the non infected feaces by the non infected population, where the formula is given here below. 
$$
rate \left[ \frac{g}{cap.day} \right] \times 365 \left[ \frac{days}{year} \right] \times (1 - \text{incidence}) \times X \left[ \text{cap} \right] = \text{production} \left[ g/year \right]
$$
On the other hand, we have the production of the non infected feaces by the infected population, during the time of the year when they are NOT infected. This is given through the formula:
$$
\text{rate} \left[ \frac{g}{cap.day} \right] \times (365 - \text{duration excretion} [\text{ days}]) = \text{production} \left[ g/year \right]
$$
The total production of non infected feaces is therefore the sum of the two. 
```{r}
production_non_infected_feaces_by_not_infected_pop <- rate * 365 * (1 - df["incidence_rate"]) * X

production_non_infected_feaces_by_infected_pop <- rate * (365- df["duration_excretion"]) * (df["incidence_rate"]) * X

df['production_non_infected_feaces'] <- production_non_infected_feaces_by_not_infected_pop + production_non_infected_feaces_by_infected_pop
```

Production of infected feaces by the part of the population that is infected is given through the equation:
$$
\text{rate} \left[ \frac{g}{cap.day} \right] \times \text{duration excretion}\left[ days \right] \times \text{incidence rate}[\text{ }] \times \text{population} = \text{production} \left[ g/year \right]
$$
In the code, this is giving:
```{r}
df['production_infected_feaces'] <- rate * df["duration_excretion"] * (df["incidence_rate"]) * X
```


## Calculation of number and density of microorganism

The number of Microorganisms is calculated with help of the excretion density of the pathogens
$$
\text{infected production} \left[\frac{g}{year} \right]*  \text{excretion density} \left[\frac{nb.org}{g} \right] = \text{Nb organisms} \left[\frac{nb. org}{year} \right] 
$$

```{r}
df['nr_pathogens_total'] <- df['production_infected_feaces']*df['excretion_density']
```

The density of microorganisms ins the according number devided by the total number of feaces produced [Nr pathogens / g of feaces]. It is calculated by the formula:
$$
\text{average density} = \frac{\text{nb.org excreted in 1 year}}{\text{total infected feces}} \quad \left[ \frac{\text{nb.org excreted}}{\text{g feces in mixed pop}} \right]
$$

```{r}
df['average_density'] <- df['nr_pathogens_total'] / (df['production_non_infected_feaces']+df['production_infected_feaces'])
print(df)
```


## Exposure pathway

Definition of parameters:
* t_hm: Transfer rate hand-mouth
* t_fh: Transfer rate fomite-hand
* t_gf: Transfer rate gloves-fomite
* t_exg: Transfer rate excreta-fomite
* a_gex: Area in contact between gloves and excreta
```{r}
t_hm <- 0.36 #[]
t_fh <- 0.43 #[]
t_gf <- 0.27 #[]
t_exg <- (0.16+0.28)/2 #[mg/cm^2]
a_gex <- (0.13+0.25)/2*420 #[cm^2]
```

We are using the equation we have seen in the Methods part:
$$
C_{Mf} = T_{H\rightarrow M} \times T_{F\rightarrow H} \times T_{G\rightarrow F} \times A_{GF} \times T_{Ex\rightarrow G} \times A_{GEx} \times C_{Ex}
$$

We first calculate the pathogen loss across the pathway, and combine it then with the densities. 
```{r}
ingestion_g <- t_exg*a_gex*t_fh*t_hm*t_gf
```

As mentionned in the study of Julian et al., 2018, there was only a small part of the workers touching their mouth. This factor can be included in order to calculate the general risk for workers working with human fertilizer. This factor can be added if we want to asses the risk for an ensemble of worker, and not just one worker touching his/her mouth. 
```{r}
f_hm <- (0.14+0.2)/2
```

Calculation of number of pathogens ingested. We want to have the result in [nb. org.  / mg feces in mixed population], therefore we need to divide by 1000
```{r}
df['number_pathogens_ingested'] <- ingestion_g*df['average_density'] /1000
 # *f_hm
print(df)
```


## Dose response models

For the dose response models, we work mainly with the exponential model and the beta-poisson model.

**Beta-poisson model:**\n
The equation used for the Beta-poisson model is the following:
$$
1-\left(1+\frac{dose}{\beta}\right)^{-\alpha}
$$
This was impemented in the code as following:
```{r}

## seems to not be okay, the one for the rota virus
response_beta_poisson_rota = 1-(1+subset(df, rownames(df) == "rota_triangular", select = "number_pathogens_ingested")/(0.42))^(-0.26)
cat("Risk of infection by rotavirus, with mode of triangular distribution:", response_beta_poisson_rota[[1]], "\n")

response_beta_poisson_rota = 1-(1+subset(df, rownames(df) == "rota_normal", select = "number_pathogens_ingested")/(0.42))^(-0.26)
cat("Risk of infection by rotavirus, with mean of normal distribution:", response_beta_poisson_rota[[1]], "\n")

response_beta_poisson_campylo1996 = 1-(1+subset(df, rownames(df) == "campylo", select = "number_pathogens_ingested")/(7.59))^(-0.145)
cat("Risk of infection by Campylobacter, using the parameters the study of 1996:", response_beta_poisson_campylo1996[[1]], "\n")

response_beta_poisson_EHEC <- 1-(1+subset(df, rownames(df) == "EHEC", select = "number_pathogens_ingested")/(39.71))^(-0.373)
cat("Risk of infection by EHEC:", response_beta_poisson_EHEC[[1]], "\n")

# Salomenlla


```

**Exponential model:**\n
The equation used for the exponential model was the following:
$$
1 - \exp(-r \times \text{dose})
$$
This was implemented in the code as following:
```{r}
response_exponential_adeno <- 1 - exp(-4.172 * 10^(-1) * subset(df, rownames(df) == "adeno", select = "number_pathogens_ingested"))
cat("Risk of infection by adenovirus:", response_exponential_adeno[[1]], "\n")

response_exponential_giardia <- 1-exp(-0.0198*subset(df, rownames(df) == "giardia", select = "number_pathogens_ingested")) 
cat("Risk of infection by giardia:", response_exponential_giardia[[1]], "\n")

response_exponential_campylo2005 = 1 - exp(-0.69 * subset(df, rownames(df) == "campylo", select = "number_pathogens_ingested"))
cat("Risk of infection by campylo, using the parameters of the study of 2005:", response_exponential_campylo2005[[1]], "\n")

#response_exponential_noro = 1-exp(-0.66 * subset(df, rownames(df) == "noro", select = "number_pathogens_ingested"))
#cat("Risk of infection by norovirus:", response_exponential_noro[[1]], "\n")


##noro


response_exponential_crypto_combined_triangular = 1-exp(-0.4*subset(df, rownames(df) == "crypto_triangular", select = "number_pathogens_ingested")) 
cat("Risk of infection by cryptosporidium parvum combined, from triangular distribution:", response_exponential_crypto_combined_triangular[[1]], "\n")

response_exponential_crypto_combined_normal = 1-exp(-0.4*subset(df, rownames(df) == "crypto_normal", select = "number_pathogens_ingested")) 
cat("Risk of infection by cryptosporidium parvum combined, from normal distribution:", response_exponential_crypto_combined_normal[[1]], "\n")

response_exponential_crypto_parvum_iowa_triangular = 1-exp(-0.0042*subset(df, rownames(df) == "crypto_triangular", select = "number_pathogens_ingested")) 
cat("Risk of infection by cryptosporidium parvum iowa, from triangular distribution:", response_exponential_crypto_parvum_iowa_triangular[[1]], "\n")

response_exponential_crypto_parvum_iowa_normal = 1-exp(-0.0042*subset(df, rownames(df) == "crypto_normal", select = "number_pathogens_ingested")) 
cat("Risk of infection by cryptosporidium parvum iowa, from normal distribution:", response_exponential_crypto_parvum_iowa_normal[[1]], "\n")

response_exponential_crypto_hominis_triangular= 1-exp(-0.76*subset(df, rownames(df) == "crypto_triangular", select = "number_pathogens_ingested")) 
cat("Risk of infection by cryptosporidium hominis, from triangular distribution:", response_exponential_crypto_hominis_triangular[[1]], "\n")

response_exponential_crypto_hominis_normal = 1-exp(-0.76*subset(df, rownames(df) == "crypto_normal", select = "number_pathogens_ingested")) 
cat("Risk of infection by cryptosporidium hominis, from normal distribution:", response_exponential_crypto_hominis_normal[[1]], "\n")

```

## Risk taking the decay into account

As it could be observed in our results, the final risk is way too high. The time of survival of the pathogens in human feaces is limited. Increasing the time between the excretion of the fertilizer and its application on the fields can therefore be used as a mitigation method. In order to assess its effectiveness, we take the calculation method implemented in Schönning et al., 2007. The formula used therefore is:

$$
H_i = M_{i-1} \left[1-\exp\left(-\frac{\ln(10)}{T_{90}}\right)\right]
$$

There are different inactivation rates in soil and in feaces. The mitigation measure can therefore be applied either to the soil or the feaces - mixing the excreta with soil will decay the pathogens in another pace than the the decay in the pure excreta. The values are taken from Schönning et al., 2007 

```{r}
df_decay <- data.frame(
  inactivation_feaces = c(30, 30, 20, 60, 60, 27.5, 70, 70),
  inactivation_soil = c(35, 35, 25,30, 30,30,495, 495),
  M1 = c(df["salmonella_triangular", "number_pathogens_ingested"], df["salmonella_normal", "number_pathogens_ingested"], df["EHEC", "number_pathogens_ingested"], df["rota_triangular", "number_pathogens_ingested"], df["rota_normal", "number_pathogens_ingested"], df["giardia", "number_pathogens_ingested"],df["crypto_triangular", "number_pathogens_ingested"], df["crypto_normal", "number_pathogens_ingested"]),
  row.names = c("salmonella_triangular", "salmonella_normal",  "EHEC","rota_triangular", "rota_normal", "giardia", "crypto_triangular", "crypto_normal")  
)
print(df_decay)
```

### Risk reduction with soil inactivation

With help of the formula mentionned above and the parameters for each pathogen, we can now have a look at the decay of the pathogens over time. For a first step, we will only consider the soil inactivation, assuming an inactivation of the pathogens in the soil. You can change the days you want to have a look at by changing the range in the loop. 

```{r}
df_decay["day_1"] = df_decay["M1"]*(1- exp(-log(10)/df_decay["inactivation_soil"]))
range_end = 12

for (i in 2:range_end) {
  M <- paste("day_", i, sep = "")
  M_1 <- paste ("day_", i-1, sep = "")
  df_decay[M] = df_decay[M_1]*(1- exp(-log(10)/df_decay["inactivation_soil"]))
}
print(df_decay)
```



In order to assess the risk going with this values, we recalculate with the dose-response models the different risks:
```{r}

df_risk <- data.frame(
)


for (i in 1:range_end) {
  day <- paste("day_", i, sep = "")
  df_risk["EHEC", day] = 1-(1+subset(df_decay, rownames(df_decay) == "EHEC", select = day)/(39.71))^(-0.373)
  df_risk["rota_triangular", day] = 1-(1+subset(df_decay, rownames(df_decay) == "rota_triangular", select = day)/(0.42))^(-0.26)
  df_risk["rota_normal", day] = 1-(1+subset(df_decay, rownames(df_decay) == "rota_normal", select = day)/(0.42))^(-0.26)
  df_risk["giardia", day] = 1-exp(-0.0198*subset(df_decay, rownames(df_decay) == "giardia", select = day))
  df_risk["crypto_combined_triangular", day] = 1-exp(-0.4*subset(df_decay, rownames(df_decay) == "crypto_triangular", select = day))
  df_risk["crypto_combined_normal", day] = 1-exp(-0.4*subset(df_decay, rownames(df_decay) == "crypto_normal", select = day)) 
  df_risk["crypto_parvum_iowa_triangular", day] = 1-exp(-0.4*subset(df_decay, rownames(df_decay) == "crypto_triangular", select = day))
  df_risk["crypto_parvum_iowa_normal", day] = 1-exp(-0.4*subset(df_decay, rownames(df_decay) == "crypto_normal", select = day))
  df_risk["crypto_hominis_triangular", day] = 1-exp(-0.4*subset(df_decay, rownames(df_decay) == "crypto_triangular", select = day))
  df_risk["crypto_hominis_normal", day] = 1-exp(-0.4*subset(df_decay, rownames(df_decay) == "crypto_normal", select = day))
}
print(df_risk)

```

### Risk reduction with feaces inactivation
The simpler inactivation would be the inactivation in the feaces itself. You can just retain the feaces for a certian period of time, and let the pathogen decay itself.
This section aims to determine the effectiveness of the approach and estimate the time required for the risk to become negligible. 

In a first step, we again calculate the number of pathogens still present with the time passing.
```{r}
df_decay["day_1"] = df_decay["M1"]*(1- exp(-log(10)/df_decay["inactivation_feaces"]))
range_end = 12

for (i in 2:range_end) {
  M <- paste("day_", i, sep = "")
  M_1 <- paste ("day_", i-1, sep = "")
  df_decay[M] = df_decay[M_1]*(1- exp(-log(10)/df_decay["inactivation_feaces"]))
}
print(df_decay)
```

The evaluation of the risk is calculated again as well: 
```{r}

df_risk <- data.frame(
)

for (i in 1:range_end) {
  day <- paste("day_", i, sep = "")
  df_risk["EHEC", day] = 1-(1+subset(df_decay, rownames(df_decay) == "EHEC", select = day)/(39.71))^(-0.373)
  df_risk["rota_triangular", day] = 1-(1+subset(df_decay, rownames(df_decay) == "rota_triangular", select = day)/(0.42))^(-0.26)
  df_risk["rota_normal", day] = 1-(1+subset(df_decay, rownames(df_decay) == "rota_normal", select = day)/(0.42))^(-0.26)
  df_risk["giardia", day] = 1-exp(-0.0198*subset(df_decay, rownames(df_decay) == "giardia", select = day))
  df_risk["crypto_combined_triangular", day] = 1-exp(-0.4*subset(df_decay, rownames(df_decay) == "crypto_triangular", select = day))
  df_risk["crypto_combined_normal", day] = 1-exp(-0.4*subset(df_decay, rownames(df_decay) == "crypto_normal", select = day)) 
  df_risk["crypto_parvum_iowa_triangular", day] = 1-exp(-0.4*subset(df_decay, rownames(df_decay) == "crypto_triangular", select = day))
  df_risk["crypto_parvum_iowa_normal", day] = 1-exp(-0.4*subset(df_decay, rownames(df_decay) == "crypto_normal", select = day))
  df_risk["crypto_hominis_triangular", day] = 1-exp(-0.4*subset(df_decay, rownames(df_decay) == "crypto_triangular", select = day))
  df_risk["crypto_hominis_normal", day] = 1-exp(-0.4*subset(df_decay, rownames(df_decay) == "crypto_normal", select = day))
}

print(df_risk)
```





