---
title: "Tool for risk assessment of pathogens in human feces"
output:
  pdf_document: default
  html_document: default
---

## Initializing code and data

To start with, we prepare the document and delete all the variables that were stored before. The entire derivation of the code is described in detail in the report.
```{r}
#clear all variables :
  rm(list = ls())

```

To start, we create a table with all relevant information. Our informations were taken from different sources. The data was given in probability distribution. In order to simplify the code and to get a point estimate, we took of the distributions the mean for normal distributions and the mode for triangular distributions. 

The sources were the following:\n
- Petterson et al., 2016 (Triangular distributions)\n
- Sch\önning et al., 2006 (Normal distributions)\n
- Jahne et al., 2017 (PERT distribution)\n

In order to find the arithmetic mean, we used the following equation for the triangular distribution:
$$
arithmetic\;mean = \frac{min + mode + max}{3}
$$
For the arithmetic mean of the normal distribution, we use the formula.
$$
arithmetic\;mean\;=\;e^{\mu+\frac{\sigma^2}{2}}
$$
For the arithmetic mean of the PERT distribution, we used the formula:
$$
arithmetic\;mean = \frac{a+4m+b}{6}
$$

With this informations, we create a table with the arithmetic means of the *excretion duration*, the *excretion density* and the *incidence rate*
```{r}
df <- data.frame(
  duration_excretion = c(
    (3+7+12)/3, 
    (3+7+12)/3,
    exp(1.6 + 1.25^2/2), 
    (15+34+42)/3, 
    (10+15+50)/3, 
    exp(3.6 + 0.2^2/2), 
    (5+10+30)/3, 
    exp(2.0 + 0.85^2/2),
    exp(4.5 + 0.7^2/2),
    exp(2.1 + 0.25^2/2), 
    (13+4*28+56)/6
    ),
  excretion_density = c(
    (10^8+10^(10)+10^(12))/3, 
    (10^8+10^(10)+10^(12))/3, 
    exp(20.7 + 2.3^2/2), 
    (10^4+10^(6)+10^(10))/3, 
    (10^6+10^(7.5)+10^9)/3, 
    exp(13.8 + 2.3^2/2), 
    (10^6+10^(7)+10^9)/3, 
    exp(17.3 + 0.6^2/2), 
    exp(15.0 + 1.7^2/2), 
    exp(5.8 + 1.2^2/2), 
    (10^(7.5)+4*10^(9.75)+10^(12))/6
    ),  
  incidence_rate = c(
    0.88/100, 
    1.28/100,
    1200/100000, 
    2.4/100, 
    0.88/100, 
    500/100000, 
    1.28/100, 
    200/100000, 
    1100/100000, 
    30/100000, 
    (428*10^4 + 696*10^4+1025*10^4)/(6*299000000)
    ),  
  row.names = c(
    "adeno", 
    "rota_triangular", 
    "rota_normal", 
    "campylo", 
    "salmonella_triangular", 
    "salmonella_normal", 
    "crypto_triangular", 
    "crypto_normal", 
    "giardia", 
    "EHEC", 
    "noro"
    )  
)

print(df)
```

We define some essential other parameters:\n
- X: Fictive population size. This parameter can be adapted according to population. For higher population samples, the variability of the results will be lower.\n
- rate: This value for the excretion has been taken as a mean value from the studies *To-Figueras et al., 2000*, *Balcells Gorina, 1989* and *Schouw et al, 2002*\n
- pathogens: In this list, we store the name of all the pathogens we listed in the table above\n
```{r}
X <- 500000             # fictive population size  
rate <- 250             # [g/cap/day] 

pathogens <- colnames(df)
```


### Production of feces
In population X, $incidence \times X$ persons are infected and $(1 - incidence) \times X$ persons are not infected during the whole year. When a person is not infected, he or she does not excrete the pathogen.

Production of not infected feaces [g/year] is devided into two subparts. For once, the production of the non infected feaces by the non infected population, where the formula is given here below. 
$$
rate \left[ \frac{g}{cap.day} \right] \times 365 \left[ \frac{days}{year} \right] \times (1 - \text{incidence}) \times X \left[ \text{cap} \right] = \text{production} \left[ g/year \right]
$$
On the other hand, we have the production of the non infected feaces by the infected population, during the time of the year when they are NOT infected. This is given through the formula:
$$
\text{rate} \left[ \frac{g}{cap.day} \right] \times (365 - \text{duration excretion} [\text{ days}]) = \text{production} \left[ g/year \right]
$$
The total production of non infected feaces is the sum of the two. 
```{r}
production_non_infected_feaces_by_not_infected_pop <- rate * 365 * (1 - df["incidence_rate"]) * X
production_non_infected_feaces_by_infected_pop <- rate * (365- df["duration_excretion"]) * (df["incidence_rate"]) * X

df['production_non_infected_feaces'] <- production_non_infected_feaces_by_not_infected_pop + production_non_infected_feaces_by_infected_pop
```

Production of infected feaces by the part of the population that is infected is given through the equation:
$$
\text{rate} \left[ \frac{g}{cap.day} \right] \times \text{duration excretion}\left[ days \right] \times \text{incidence rate}[\text{ }] \times \text{population} = \text{production} \left[ g/year \right]
$$
In the code, this is giving:
```{r}
df['production_infected_feaces'] <- rate * df["duration_excretion"] * (df["incidence_rate"]) * X
```


### Calculation of number and density of microorganism
The number of Microorganisms is calculated with help of the excretion density of the pathogens
$$
\text{infected production} \left[\frac{g}{year} \right]*  \text{excretion density} \left[\frac{nb.org}{g} \right] = \text{Nb organisms} \left[\frac{nb. org}{year} \right] 
$$

```{r}
df['nr_pathogens_total'] <- df['production_infected_feaces']*df['excretion_density']
```

The density of microorganisms ins the according number devided by the total number of feaces produced [Nr pathogens / g of feaces]. It is calculated by the formula:
$$
\text{average density} = \frac{\text{nb.org excreted in 1 year}}{\text{total infected feces}} \quad \left[ \frac{\text{nb.org excreted}}{\text{g feces in mixed pop}} \right]
$$

```{r}
df['total_feces_produced'] <- (df['production_non_infected_feaces']+df['production_infected_feaces'])
df['average_density'] <- df['nr_pathogens_total'] / df['total_feces_produced']
```

## Exposure pathway

Definition of parameters:\n
- t_hm: Transfer rate hand-mouth\n
- t_fh: Transfer rate fomite-hand\n
- t_gf: Transfer rate gloves-fomite\n
- t_exg: Transfer rate excreta-fomite\n
- a_gex: Area in contact between gloves and excreta\n
```{r}
t_hm <- 0.36 #[]
t_fh <- 0.43 #[]
t_gf <- 0.27 #[]
t_exg <- (0.16+0.28)/2 #[mg/cm^2]
a_gex <- (0.13+0.25)/2*420 #[cm^2]
```

We are using the equation we have seen in the Methods part:
$$
C_{Mf} = T_{H\rightarrow M} \times T_{F\rightarrow H} \times T_{G\rightarrow F} \times A_{GF} \times T_{Ex\rightarrow G} \times A_{GEx} \times C_{Ex}
$$

We first calculate the pathogen loss across the pathway, and combine it then with the densities. 
```{r}
ingestion_g <- t_exg*a_gex*t_fh*t_hm*t_gf
```

After having calculated the exposure pathway, we want to apply this on the number of pathogens ingested. There are two particularities of the data, that are adapted in this part of the code:\n
- We want to have the result in [nb. org.  / mg feces in mixed population], therefore we need to divide the average pathogen density by 1000\n
- For the rota virus, we need the dose in the unit [FFU]. The transformation between nr of pathogens and FFU is given by the multiplication by 1900. $$dose\;[nb. org] = dose\;[FFU]*1900$$\n
```{r}
df['number_pathogens_ingested'] <- ingestion_g*df['average_density'] /1000

# Multiply the specified cell by 1900 => we need to do this, in order to have the right unit for the dose response model
df["rota_triangular", "number_pathogens_ingested"] <- df["rota_triangular", "number_pathogens_ingested"] *1900
df["rota_normal", "number_pathogens_ingested"] <- df["rota_normal", "number_pathogens_ingested"] * 1900

```

## Dose response models
For the dose response models, we work mainly with the exponential model and the beta-poisson model.

**Beta-poisson model:**
The equation used for the Beta-poisson model is the following:
$$
1-\left(1+\frac{dose}{\beta}\right)^{-\alpha}
$$

**Exponential model:**\n
The equation used for the exponential model was the following:
$$
1 - \exp(-r \times \text{dose})
$$
Those models were implemented with the according parameters given in the report in the following section. The results were then saved in a dataframe, in order to have it all in a more compact way. 
```{r}
response_beta_poisson_campylo1996 = 1-(1+subset(df, rownames(df) == "campylo", select = "number_pathogens_ingested")/(7.59))^(-0.145)
response_beta_poisson_EHEC <- 1-(1+subset(df, rownames(df) == "EHEC", select = "number_pathogens_ingested")/(39.71))^(-0.373)

response_exponential_salmonella_triangular = 1 - exp(-3.97 * 10^(-6) * subset(df, rownames(df) == "salmonella_triangular", select = "number_pathogens_ingested"))
response_exponential_salmonella_normal = 1 - exp(-3.97 * 10^(-6) * subset(df, rownames(df) == "salmonella_normal", select = "number_pathogens_ingested"))
response_exponential_adeno = 1 - exp(-4.172 * 10^(-1) * subset(df, rownames(df) == "adeno", select = "number_pathogens_ingested"))
response_exponential_giardia = 1-exp(-0.0198*subset(df, rownames(df) == "giardia", select = "number_pathogens_ingested")) 
response_exponential_campylo2005 = 1 - exp(-0.69 * subset(df, rownames(df) == "campylo", select = "number_pathogens_ingested"))
response_exponential_noro = 1-exp(-0.66 * subset(df, rownames(df) == "noro", select = "number_pathogens_ingested"))
response_exponential_rota_normal = 1-exp(-0.47 * subset(df, rownames(df) == "rota_normal", select = "number_pathogens_ingested"))
response_exponential_rota_triangular = 1-exp(-0.47 * subset(df, rownames(df) == "rota_triangular", select = "number_pathogens_ingested"))
response_exponential_crypto_combined_triangular = 1-exp(-0.4*subset(df, rownames(df) == "crypto_triangular", select = "number_pathogens_ingested")) 
response_exponential_crypto_combined_normal = 1-exp(-0.4*subset(df, rownames(df) == "crypto_normal", select = "number_pathogens_ingested")) 
response_exponential_crypto_parvum_iowa_triangular = 1-exp(-0.0042*subset(df, rownames(df) == "crypto_triangular", select = "number_pathogens_ingested")) 
response_exponential_crypto_parvum_iowa_normal = 1-exp(-0.0042*subset(df, rownames(df) == "crypto_normal", select = "number_pathogens_ingested")) 
response_exponential_crypto_hominis_triangular= 1-exp(-0.76*subset(df, rownames(df) == "crypto_triangular", select = "number_pathogens_ingested")) 
response_exponential_crypto_hominis_normal = 1-exp(-0.76*subset(df, rownames(df) == "crypto_normal", select = "number_pathogens_ingested"))
```

```{r}
df_responses <- data.frame(
  responses_daily = c(response_beta_poisson_campylo1996[[1]],
                response_beta_poisson_EHEC[[1]],
                response_exponential_salmonella_triangular[[1]],
                response_exponential_salmonella_normal[[1]],
                response_exponential_adeno[[1]], 
                response_exponential_giardia[[1]], 
                response_exponential_campylo2005[[1]], 
                response_exponential_noro[[1]], 
                response_exponential_rota_normal[[1]], 
                response_exponential_rota_triangular[[1]], 
                response_exponential_crypto_combined_triangular[[1]], 
                response_exponential_crypto_combined_normal[[1]], 
                response_exponential_crypto_parvum_iowa_triangular[[1]], 
                response_exponential_crypto_parvum_iowa_normal[[1]], 
                response_exponential_crypto_hominis_triangular[[1]], 
                response_exponential_crypto_hominis_normal[[1]]),
  concentrations = c(df['campylo', "number_pathogens_ingested"],
                df['EHEC', "number_pathogens_ingested"],
                df['salmonella_triangular', "number_pathogens_ingested"],
                df['salmonella_normal', "number_pathogens_ingested"],
                df['adeno', "number_pathogens_ingested"],
                df['giardia', "number_pathogens_ingested"],
                df['campylo', "number_pathogens_ingested"],
                df['noro', "number_pathogens_ingested"],
                df['rota_normal', "number_pathogens_ingested"],
                df['rota_triangular', "number_pathogens_ingested"],
                df['crypto_triangular', "number_pathogens_ingested"],
                df['crypto_normal', "number_pathogens_ingested"],
                df['crypto_triangular', "number_pathogens_ingested"],
                df['crypto_normal', "number_pathogens_ingested"],
                df['crypto_triangular', "number_pathogens_ingested"],
                df['crypto_normal', "number_pathogens_ingested"]
                ),
  row.names =  c("response_exponential_campylo1996",
                 "response_exponential_EHEC",
                 "response_exponential_salmonella_triangular",
                 "response_exponential_salmonella_normal",
                 "response_exponential_adeno",
                "response_exponential_giardia",
                "response_exponential_campylo2005",
                "response_exponential_noro",
                "response_exponential_rota_normal",
                "response_exponential_rota_triangular",
                "response_exponential_crypto_combined_triangular",
                "response_exponential_crypto_combined_normal",
                "response_exponential_crypto_parvum_iowa_triangular",
                "response_exponential_crypto_parvum_iowa_normal",
                "response_exponential_crypto_hominis_triangular",
                "response_exponential_crypto_hominis_normal")
)
print(df_responses)
```

## Log reduction of risk
The U.S. *Environmental Protection Agency* propose a benchmark for the  the acceptable annual infection risk level of of $10^{-4}$ infections $person^{-1}$ $year^{-1}$.

Thought the exposure pathoway, the point estimation of probability of infection based on a one-time pathogen exposure was computed for each pathogen. In order to compare those results to the U.S. *Environmental Protection Agency* threshold, the probability of infections need to converted on an annual basis. The same procedure is used than in the study of Brooks et al., 2012.

$$
P_{ann} = 1-(1-P_i)^n
$$
$P_{ann}$ is the probability of infection based on n, the number of days or exposure events per year. $P_i$ is the probability of infection on a one-time pathogen exposure. 

In \cite{brooks_land_2012}, they defined occupational annual exposures as n = 255 d/yr (USEPA, 1997), which correspond approximately to 52 week/year x 5 working days / week.  However, in Switzerland, solid manure is often spread once or twice a year. From these considerations, the value of n=2 was chosen.

$$
P_{ann} = 1-(1-P_i)^{2}
$$

This was implemented in the code as following:
```{r}
df_responses["response_yearly"] = 1-(1-df_responses["responses_daily"])^2
print(df_responses)
```

In order to know the log reduction we need to have an acceptable risk, we assess the risk for different levels of log reductions. The final risk must be beneath $10^{-4}$ infections $person^{-1}$ $year^{-1}$., as shown above. Whe therefore create a dataframe having the concentrations for the different log reductions and then a second one displaying the according risks.
```{r}
df_log <- data.frame(log_0 = c(df['adeno', "average_density"], 
                               df['rota_triangular', "average_density"], 
                               df['rota_normal', "average_density"], 
                               df['campylo', "average_density"], 
                               df['salmonella_triangular', "average_density"], 
                               df['salmonella_normal', "average_density"],
                               df['EHEC', "average_density"],
                               df['giardia', "average_density"],
                               df['crypto_triangular', "average_density"],
                               df['crypto_normal', "average_density"],
                               df['noro', "average_density"]),
  row.names = c("adeno", 
    "rota_triangular", 
    "rota_normal", 
    "campylo", 
    "salmonella_triangular", 
    "salmonella_normal", 
    "EHEC",
    "giardia",
    "crypto_triangular", 
    "crypto_normal", 
    "noro")
)


range_end = 14

for (i in 1:range_end) {
  M <- paste("log_", i, sep = "")
  M_1 <- paste ("log_", i-1, sep = "")
  df_log[M] = df_log[M_1]/10
}

df_risk <- data.frame(
)

f <- function(x) {
  1 - (1 - x)^(2)
}


for (i in 1:range_end) {
  log <- paste("log_", i, sep = "")
  df_risk["EHEC", log] = sapply(1-(1+subset(df_log, rownames(df_log) == "EHEC", select = log)/(39.71))^(-0.373), f)
  df_risk["salmonella_triangular", log] = sapply(1 - exp(-3.97 * 10^(-6) * subset(df_log, rownames(df_log) == "salmonella_triangular", select = log)), f)
  df_risk["salmonella_normal", log] = sapply(1 - exp(-3.97 * 10^(-6) * subset(df_log, rownames(df_log) == "salmonella_normal", select = log)), f)
  df_risk["campylo_2005", log] = sapply(1 - exp(-0.69 * subset(df_log, rownames(df_log) == "campylo", select = log)), f)
  df_risk["campylo_1996", log] =  sapply(1-(1+subset(df_log, rownames(df_log) == "campylo", select = log)/(7.59))^(-0.145), f)
  df_risk["giardia", log] =  sapply(1-exp(-0.0198*subset(df_log, rownames(df_log) == "giardia", select = log)), f)
  df_risk["adeno", log] =  sapply(1 - exp(-4.172 * 10^(-1) * subset(df_log, rownames(df_log) == "adeno", select = log)), f)
  df_risk["noro", log] = sapply(1-exp(-0.66 * subset(df_log, rownames(df_log) == "noro", select = log)), f)
  df_risk["rota_triangular", log] = sapply(1-(1+subset(df_log, rownames(df_log) == "rota_triangular", select = log)/(0.42))^(-0.26), f)
  df_risk["rota_normal", log] = sapply(1-(1+subset(df_log, rownames(df_log) == "rota_normal", select = log)/(0.42))^(-0.26), f)
  df_risk["giardia", log] = sapply(1-exp(-0.0198*subset(df_log, rownames(df_log) == "giardia", select = log)), f)
  df_risk["crypto_combined_triangular", log] = sapply(1-exp(-0.4*subset(df_log, rownames(df_log) == "crypto_triangular", select = log)), f)
  df_risk["crypto_combined_normal", log] = sapply(1-exp(-0.4*subset(df_log, rownames(df_log) == "crypto_normal", select = log)) , f)
  df_risk["crypto_parvum_iowa_triangular", log] = sapply(1-exp(-0.0042*subset(df_log, rownames(df_log) == "crypto_triangular", select = log)), f)
  df_risk["crypto_parvum_iowa_normal", log] = sapply(1-exp(-0.0042*subset(df_log, rownames(df_log) == "crypto_normal", select = log)), f)
  df_risk["crypto_hominis_triangular", log] = sapply(1-exp(-0.76*subset(df_log, rownames(df_log) == "crypto_triangular", select = log)), f)
  df_risk["crypto_hominis_normal", log] = sapply(1-exp(-0.76*subset(df_log, rownames(df_log) == "crypto_normal", select = log)), f)
}


print(df_risk)

```

## Risk taking the decay into account

As it could be observed in our results, the final risk is way too high. The time of survival of the pathogens in human feaces is limited. Increasing the time between the excretion of the fertilizer and its application on the fields can therefore be used as a mitigation method. In order to assess its effectiveness, we take the calculation method implemented in Schönning et al., 2007. The formula used therefore is:

$$
H_i = M_{i-1} \left[1-\exp\left(-\frac{\ln(10)}{T_{90}}\right)\right]
$$

There are different inactivation rates in soil and in feaces. The mitigation measure can therefore be applied either to the soil or the feaces - mixing the excreta with soil will decay the pathogens in another pace than the the decay in the pure excreta. The values are taken from Schönning et al., 2007 

```{r}
df_decay <- data.frame(
  inactivation_feaces = c(30, 30, 20, 60, 60, 27.5, 70, 70),
  M1 = c(df["salmonella_triangular", "nr_pathogens_total"], 
         df["salmonella_normal", "nr_pathogens_total"], 
         df["EHEC", "nr_pathogens_total"], 
         df["rota_triangular", "nr_pathogens_total"], 
         df["rota_normal", "nr_pathogens_total"], 
         df["giardia", "nr_pathogens_total"],
         df["crypto_triangular", "nr_pathogens_total"], 
         df["crypto_normal", "nr_pathogens_total"]),
  total_feces = c(df["salmonella_triangular", "total_feces_produced"], 
                  df["salmonella_normal", "total_feces_produced"], 
                  df["EHEC", "total_feces_produced"], 
                  df["rota_triangular", "total_feces_produced"], 
                  df["rota_normal", "total_feces_produced"], 
                  df["giardia", "total_feces_produced"],
                  df["crypto_triangular", "total_feces_produced"], 
                  df["crypto_normal", "total_feces_produced"]),
  row.names = c("salmonella_triangular", "salmonella_normal",  "EHEC","rota_triangular", "rota_normal", "giardia", "crypto_triangular", "crypto_normal")  
)
print(df_decay)
```


## Risk reduction with feaces inactivation
In order to achieve this log reduction, we somehow need to reduce the concentration of pathogens in the feces. A simple mitigation method would be the retention of the feces for a certain time, alloing the pathogens to decay on themselves.This section aims to determine the effectiveness of the approach and estimate the time required for the risk to become negligible. 

In a first step, we again calculate the number of pathogens still present with the time passing. We calculate the decay on the *total number of pathogens*, and apply then again the formulas we saw above for the *average density* and the *number of pathogens ingested*
```{r}
df_decay["day_0"] = df_decay["M1"]
df_decay["day_1"] = df_decay["M1"]*(1- exp(-log(10)/df_decay["inactivation_feaces"]))
range_end = 12

for (i in 2:range_end) {
  M <- paste("day_", i, sep = "")
  M_1 <- paste ("day_", i-1, sep = "")
  df_decay[M] = df_decay[M_1]*(1- exp(-log(10)/df_decay["inactivation_feaces"]))
}

```

```{r}
df_average_density <- df_decay / df[["total_feces_produced"]]
df_average_density <- subset(df_average_density, select = -c(M1, inactivation_feaces, total_feces))

df_nr_path_ingested <- ingestion_g*df_average_density/1000

print(df_nr_path_ingested)

```

We again apply the function $$P_{ann} = 1-(1-P_i)^{255}$$ used to calculate yearly risk values.  
```{r}
f <- function(x) {
  1 - (1 - x)^(2)
}

df_risk_decay <- data.frame(
)

for (i in 1:range_end) {
  day <- paste("day_", i, sep = "")
  df_risk_decay["EHEC", day] = sapply(1-(1+subset(df_nr_path_ingested, rownames(df_nr_path_ingested) == "EHEC", select = day)/(39.71))^(-0.373), f)
  df_risk_decay["salmonella_triangular", day] = sapply(1-exp(-3.97*10^(-6)*subset(df_nr_path_ingested, rownames(df_nr_path_ingested)=="salmonella_triangular", select = day)), f)
  df_risk_decay["salmonella_normal", day] = sapply(1 - exp(-3.97 * 10^(-6) * subset(df_nr_path_ingested, rownames(df_nr_path_ingested) == "salmonella_normal", select = day)), f)
  df_risk_decay["rota_triangular", day] = sapply(1-(1+subset(df_nr_path_ingested, rownames(df_nr_path_ingested) == "rota_triangular", select = day)/(0.42))^(-0.26), f)
  df_risk_decay["rota_normal", day] = sapply(1-(1+subset(df_nr_path_ingested, rownames(df_nr_path_ingested) == "rota_normal", select = day)/(0.42))^(-0.26), f)
  df_risk_decay["giardia", day] = sapply(1-exp(-0.0198*subset(df_nr_path_ingested, rownames(df_nr_path_ingested) == "giardia", select = day)), f)
  df_risk_decay["crypto_combined_triangular", day] = sapply(1-exp(-0.4*subset(df_nr_path_ingested, rownames(df_nr_path_ingested) == "crypto_triangular", select = day)), f)
  df_risk_decay["crypto_combined_normal", day] = sapply(1-exp(-0.4*subset(df_nr_path_ingested, rownames(df_nr_path_ingested) == "crypto_normal", select = day)) , f)
  df_risk_decay["crypto_parvum_iowa_triangular", day] = sapply(1-exp(-0.0042*subset(df_nr_path_ingested, rownames(df_nr_path_ingested) == "crypto_triangular", select = day)), f)
  df_risk_decay["crypto_parvum_iowa_normal", day] = sapply(1-exp(-0.0042*subset(df_nr_path_ingested, rownames(df_nr_path_ingested) == "crypto_normal", select = day)), f)
  df_risk_decay["crypto_hominis_triangular", day] = sapply(1-exp(-0.76*subset(df_nr_path_ingested, rownames(df_nr_path_ingested) == "crypto_triangular", select = day)), f)
  df_risk_decay["crypto_hominis_normal", day] = sapply(1-exp(-0.76*subset(df_nr_path_ingested, rownames(df_nr_path_ingested) == "crypto_normal", select = day)), f)
}

print(df_risk_decay)
```
